tstrait: a quantitative trait simulator for ancestral recombination graphs

### 2023-11-07 paper.tex
Update

Genome-wide association studies (GWAS)
identify genetic variants that are statistically
associated with a specific trait~\citep{uffelmann2021}.
GWAS has successfully
identified many loci that are associated with various human diseases and
traits~\citep{locke2015,ishigaki2022,mahajan2022,yengo2022,mathieson2023},
and GWAS results are being actively incorporated into clinical
practice~\citep{visscher2017}.
Despite these successes, however, there are fundamental issues with
In spite of its advantages, population structure and
demographic histories can heavily influence GWAS results \citep{uffelmann2021},
so it would be necessary for researchers to implement these factors when
conducting GWAS.

Ancestral recombination graphs (ARGs) have many promising aspects in GWAS.
Recent studies have shown that genealogy-based association can detect more
ultra rare variants than conventional methods \citep{zhang2023}, using an
ARG-based approach in quantitative-trait locus mapping can better detect causal
locus, especially in the presence of allelic heterogeneity \citep{link2023},
and linkage disequilibirium can be efficiently modeled in GWAS by using a
tree-based approach \citep{salehi2023}. These recent advances make ARG to be a
suitable data format for conducting GWAS.

>>>>

Genome-wide association studies (GWAS)
identify genetic variants that are statistically
associated with a specific trait~\citep{uffelmann2021}.
GWAS has successfully
identified many loci that are associated with various human diseases and
traits~\citep{locke2015,ishigaki2022,mahajan2022,yengo2022,mathieson2023},
and GWAS results are actively being incorporated into clinical
practice~\citep{visscher2017}.
Despite these successes and the ongoing collection of population-scale
Biobank data~\citep{tanjo2021practical}, there are substantial open
questions around GWAS methology.
In particular, correctly accounting for population stratification
remains challenging~\citep{uffelmann2021}.

Ancestral recombination graphs (ARGs) describe the interwoven paths
of genetic inheritance arising from recombination~\citep{wong2023general}.
[TODO from here]
Recent studies have shown that genealogy-based association can detect more
ultra rare variants than conventional methods \citep{zhang2023}, using an
ARG-based approach in quantitative-trait locus mapping can better detect causal
locus, especially in the presence of allelic heterogeneity \citep{link2023},
and linkage disequilibirium can be efficiently modeled in GWAS by using a
tree-based approach \citep{salehi2023}. These recent advances make ARG to be a
suitable data format for conducting GWAS.
